---
title: "week_04_notes"
author: "Andrew Nalundasan"
date: "7/9/2021"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggplot2)
library(vtable)
library(grid)
library(gridExtra)
library(lattice)
library(janitor)
library(GGally)
```

# Overview Video

+ Focus on getting good at ggplot2

+ Data Vizualization checklist

    - URL
    
+ pie charts

    - limited value
    - difficult to tell relative sizes of different categories
    - works fine with only 2 categories
    - 3 categories is still difficult to tell a good story
    
+ tree map is a good alternative to pie chart

    - instead of a pie, it's squares and blocks
    - library(treemapify)
    - geom_treemap()
   
+ scales functions

    - set limits of the graph
    - bar graph: 
        - limits = c("C", "A", "B")
        - this makes the bars in C A B order

+ labels

    - can use unnamed functions
    - function(x) or \(x)
    
+ theme

    - strip 
        - useful when using facets to label the different facets
    - title
    - element
        - element text, element rectangles, element blank, etc.
    - geometry
        - if aesthetic is INSIDE the aesthetic function, it applies to just the one element
        - if aesthetic is OUTSIDE the aesthetic function, it applies to the whole viz
    - font size is determined differently between theme vs geometry
    - theme is pretty standard how we know it
    - geometry you must divide by .pt
        - size = 12/.pt
        - this gives standard 12 point font
        - conversion is also necessary in annotate() functions

+ guides
    
    - color = 'none' to remove legends
    
```{r}
# knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

# this will remove messages and warnings from R outputs. 
```


+ annotation

    - prefer annotation over legends (where possible)
        - move the annotation closer to the data
    - theme(legend.position) can move the legend closer to the data
    
    
# Best Practices

+ Principles are straightforward but application is often missed 

    - most time spent on analysis
    - need time to spend on communication
    
+ Keep it simple

    - 1 main point
    - maybe 1-2 sub-points
    
+ Use the appropriate graph

    - time-series or continuation needs a line graph rather than bar graph
    
+ Explanatory visuals have appropriate titles and labels

    - expert graph readers look at: 
        - title > axes > data
    - novice graph readers are all over the place
        - tell these people what they need to see
    - 5 or less variables is best, 8 variables max
        
        
+ Be intentional with colors

+ Make data readable

    - axes and labels must be readable
    
+ Double check your numbers

+ Bad visualizations come from 3 sources

    - Aesthetic
        - inconsistent, ugly, random
    - Substantive
        - the data is bad
        - ensure you're graphing the right data
    - Perceptual
        - confusing or misleading
        - difficult for audience to see/process what they're looking at
        
## Keep it Simple

+ What's the point of the graph?

    - Remove clutter
    - Ensure proper axis labels
    
+ Use the appropriate graph

    - Line graph: shows CONTINUATION, shows MOTION
    - Bar graph: shows quantities and no continuation
        - Make certain origin is at 0
    - Pie charts: the eye cannot determine the size of angular shapes
        - 2 variables is the best for pie chart to determine similarity or differences
        - Show name and % on each slice
    - Always better to not have a legend
    
+ Appropriate titles and labels

    - Graph title indicates the main point trying to make
    - Titles and labels gets audience to focus attention where you want
    - Ensure scale is uniform
    - Flip coordinates before making your audience turn their heads to read labels
    
+ Be careful with colors

    - Humans can only distinguis 8 colors at a time
        - 5 colors max is best practice
    
+ Make data readable

    - 3D is a bad idea
    - Be intentional with scale
        - "This vs. everything else"
        
+ Typography / Iconography

    - Font can communicate meaning
    - Floppy Disk <- save
    

# Clutter is your enemy

+ Make graphs look more professional

+ None of the analysis or modeling matters if it is not communicated well

+ Often, people leave little time to build the communication and visuals

## Cognitive overload and clutter

+ Don't try to do too much. Do less. 

+ Goal: Make life easy for the reader

    - Do the work for your audience
    - Make your work easy to understand
    - Know what message you are trying to communicate
    - Remove your clutter

+ Written examples:
  
    - Bullet format is easier to process
         - "To the point"

+ Clutter:

    - Extra visual elements that take up space but don't increase understanding
    - "I didn't have time to write a short letter, so I wrote a long one instead."
    - Maximize the signal-to-noise ratio
        - aka "ink ratio"
        - make you audience able to FOCUS on the message
    - No cognitive overloade; reduce the clutter
    

# Gestalt principles

+ Reducing clutter <- grouping information naturally

    - Ability to maintain meaningful perceptions
    - How people perceive order
    
+ Gestalt Principles

    - Similarity - items that look the same belong together
    - Proximity - items closer together belong together
    - Enclosure - items enclosed together are part of a group
    - Closure - people perceive parts as part of a whole
        - people group things naturally
        - borders usually unnecessary
        - gridlines usually unnecessary
    - Continuity - people look for the smootheset path
        - people expect things to be continuous
    - Connection - people think of items that are connected as part of a group
    
+ Using Gestalt principles

    - Declutter visuals because of people's natural abilities
    - Decluttering is an art; data viz is an art


# Visual order, part 1

+ Helps people understand your main point

    - If graph has order, people will follow naturally what you're trying to say
    
    
## Orientation

+ Sometimes easier to read a graph with flipped axes

    - ggplot's default organizes alphabetically
    
```{r}
# load data
dis <- read_csv("Disability.csv")

# Recode date to age range, this is to fix an Excel formatting problem
dis <- dis %>% 
  mutate(age_group = recode(age_group, '12-Jun' = "6-12"))
```
    
    
```{r}
# Default graphic of median expenditures by ethnicity
dis %>% 
  group_by(ethnicity) %>% 
  summarize(med_exp = median(expenditures)) %>% 
  ggplot(mapping = aes(x = ethnicity, y = med_exp)) + 
  geom_bar(stat = "identity")
```


```{r}
# change orientation
dis %>% 
  group_by(ethnicity) %>% 
  summarize(med_exp = median(expenditures)) %>% 
  ggplot(mapping = aes(x = ethnicity, y = med_exp)) + 
  geom_bar(stat = "identity") + 
  coord_flip()

```


### Reorder by numeric variable

+ Order bars by median expenditures rather than alphabetically

```{r}
# Order bars by median expenditures
dis %>% 
  group_by(ethnicity) %>% 
  summarize(med_exp = median(expenditures)) %>% 
  ggplot(mapping = aes(x = reorder(ethnicity, med_exp), y = med_exp)) + 
  geom_bar(stat = "identity") + 
  coord_flip()
```

+ We can reorder factor variables based on any quantitative variable in this manner

### Reorder by *other* numeric variable

```{r}
# Order bars on count while graphing median expenditures
dis %>% 
  group_by(ethnicity) %>% 
  summarize(med_exp = median(expenditures),
            n_eth = n()) %>% 
  ggplot(mapping = aes(x = reorder(ethnicity, n_eth), y = med_exp)) + 
  geom_bar(stat = "identity") + 
  coord_flip()
```

+ Order by count if the need arises

### Reorder by factor levels

+ reorder the facor using "fct_relevel()" within "mutate()"

```{r}
# Order factor level with order vector
ord <- c("Other", "White", "American Indian", "Native Hawaiian", "Multi-Race", "Asian", "Hispanic", "Black")

dis %>%
  mutate(ethnicity = fct_relevel(ethnicity, levels = rev(ord))) %>%   # "rev" makes the list go from top down. default starts at origin
  group_by(ethnicity) %>% 
  summarize(med_exp = median(expenditures)) %>% 
  ggplot(mapping = aes(x = ethnicity, y = med_exp)) + 
  geom_bar(stat = "identity") + 
  coord_flip()
```


### Reorder bars within clusters

```{r}
# Order clusters by median expenditure and include gender
dis %>% 
  group_by(ethnicity, gender) %>% 
  summarize(med_exp = median(expenditures)) %>% 
  ggplot(mapping = aes(x = reorder(ethnicity, med_exp), y = med_exp, fill = gender)) + 
  geom_bar(stat = "identity", position = "dodge") + 
  coord_flip()
```

```{r}
# Order clusters and bars by median expenditure and include gender
dis %>% 
  group_by(ethnicity, gender) %>% 
  summarize(med_exp = median(expenditures)) %>% 
  ggplot(mapping = aes(x = reorder(ethnicity, med_exp), y = med_exp, fill = reorder(gender, med_exp))) +  # set fill to reorder.
  geom_bar(stat = "identity", position = "dodge") + 
  coord_flip()
```

### Reorder the legend

+ be visually consistent by reordering the legend

```{r}
# Order factor legend within code chunk without using order vector
dis %>% 
  group_by(ethnicity, gender) %>% 
  summarize(med_exp = median(expenditures)) %>% 
  ggplot(mapping = aes(x = reorder(ethnicity, med_exp), y = med_exp, fill = reorder(gender, med_exp))) + 
  geom_bar(stat = "identity", position = "dodge") + 
  coord_flip() + 
  scale_fill_discrete(limits = c("Female", "Male"))
```

### Visual Order

+ Order legend by continuous variable rather than discrete variable

+ Know WHAT you want to do, THEN figure out HOW to do it in the software


# Visual order, part 2

+ Visual order relative to: 

    - Left alignment is better than being centered
        - Titles
        - Text
    - Diagonal components should be avoided
        - Diagonal lines are distracting
        - Diagonal text is hard to read
    - Stragic use of contracts
        - Make important bars stand out
        - Be intentional with color
        - Don't be too dramatic
    - White space
        - Avoid filling the entire graph area
        - Removal of lines create white space
        - Use contrast or location of text to remove the need for arrow-lines
        
### Themes

```{r}
# Theme classic removes gray lines in background
dis %>% 
  group_by(ethnicity, gender) %>% 
  summarize(med_exp = median(expenditures)) %>% 
  ggplot(mapping = aes(x = reorder(ethnicity, med_exp), y = med_exp, fill = reorder(gender, med_exp))) + 
  geom_bar(stat = "identity", position = "dodge") + 
  coord_flip() + 
  scale_fill_discrete(limits = c("Female", "Male")) + 
  theme_classic()
```

### Remove tick marks

```{r}
# Remove lines and tick marks to reduce clutter
dis %>% 
  group_by(ethnicity, gender) %>% 
  summarize(med_exp = median(expenditures)) %>% 
  ggplot(mapping = aes(x = reorder(ethnicity, med_exp), y = med_exp, fill = reorder(gender, med_exp))) + 
  geom_bar(stat = "identity", position = "dodge") + 
  coord_flip() + 
  scale_fill_discrete(limits = c("Female", "Male")) + 
  theme_classic() + 
  theme(axis.line.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.ticks.x = element_blank())
```

### ggthemes

```{r}
# load library
library(ggthemes)

# Apply Wall Street Journal theme
# Order factor legend within code chunk without using order vector
dis %>% 
  group_by(ethnicity, gender) %>% 
  summarize(med_exp = median(expenditures)) %>% 
  ggplot(mapping = aes(x = reorder(ethnicity, med_exp), y = med_exp, fill = reorder(gender, med_exp))) + 
  geom_bar(stat = "identity", position = "dodge") + 
  coord_flip() + 
  scale_fill_discrete(limits = c("Female", "Male")) + 
  theme_wsj()
```

+ Theme function has heaps of settings

### Creating custom themes

+ start with theme_classic() and add changes

```{r}
# Start with theme_classic() and modify a few things
mytheme <- theme_classic() + 
  theme(text = element_text(colour = "red"), 
        axis.title = element_text(size = rel(1.25)), 
                                  plot.background = element_rect(fill = "light blue"))
```

```{r}
# The base graph we have been using plus mytheme
dis %>% 
  group_by(ethnicity, gender) %>% 
  summarize(med_exp = median(expenditures)) %>% 
  ggplot(mapping = aes(x = reorder(ethnicity, med_exp), y = med_exp, fill = reorder(gender, med_exp))) + 
  geom_bar(stat = "identity", position = "dodge") + 
  coord_flip() + 
  scale_fill_discrete(limits = c("Female", "Male")) + 
  mytheme
```

+ Themes are powerful

    1. Reduce clutter
        - Highlight the points you want to highlight
    2. Customize plots
    3. Save time by applying custom theme and reuse it

# Focusing audience attention

+ Ensure your audience see what you want them to see

+ EDA is all about finding the important message
    
    - Need to invest time and effort to COMMUNICATE the important finding
    - Make it easy for the audience to follow along
    
+ What is the goal, what's your plan, what are you trying to communicate?

    - Consider what your audience will see and remember
    - Preattentive Attributes!!
    
+ Preattentive Attributes
  
    - Use these to direct your audiences attention
    - Use to build a visual hierarchy to lead your audience through the information

+ Take a stand and show what's important

    - Text: color and bold are most effective
        - *Italics* and underline are not very effective
    - Title needs to stand out. This is the main idea of the visual!
    - Be intentional with color
        - Emphasize what's important
        - De-emphasize what's less important
    - Size/shape
    - Annotations
        - Enclosures to make things standout
        - Text
      

# Titles, labels, and legends

+ Informative text used to describe the visual

+ Don't let the title repeat the axes

    - Think of set of labels as a WHOLE for the visual
    
## Best Practices    

+ Title and Subtitle: indicate the main/sub points you want the audience to see in the visual
    
    - Always set while graphing
    - Avoid long titles, be concise
    
+ Axes and legend labels: variable and units (if applicable)

    - Often taken from the data, but you can format factor level names or digits while wrangling or graphing
    
+ Tick and legend labels: factor level or quantity

    - Format appropriately (no scientific notation, %, $, ",", etc)
    - Often taken directly from the data, but you can format factor level names or digits in the data or while wrangling
    
+ Reference labels: additional labels that could include a figure number or source

## Primary titles and labels

```{r}
# The base graph we have been redesigning plus labels
dis %>% 
  group_by(ethnicity, gender) %>% 
  summarize(med_exp = median(expenditures)) %>% 
  ggplot(mapping = aes(x = reorder(ethnicity, med_exp), y = med_exp, fill = reorder(gender, med_exp))) + 
  geom_bar(stat = "identity", position = "dodge") + 
  coord_flip() + 
  scale_fill_discrete(limits = c("Female", "Male")) + 
  theme_classic() + 
  labs(title = "Main title\nmain title continued",
       subtitle = "Subtitle", 
       fill = "Legend\ntitle",
       x = "New x-axis label", 
       y = "New y-axis label", 
       tag = "Tag",     # would likely be a figure number or something else
       caption = "Caption")    # would likely be a source or reference
```

```{r}
# Remake of overall graph
p <- dis %>% 
  filter(ethnicity %in% c("Asian", "Black", "Hispanic", "White")) %>% 
  group_by(ethnicity, gender) %>% 
  summarize(med_exp = median(expenditures)) %>% 
  ggplot(mapping = aes(x = reorder(ethnicity, med_exp), y = med_exp, fill = gender)) + 
  geom_bar(stat = "identity", position = "dodge") + 
  coord_flip() + 
  scale_fill_discrete(limits = c("Male", "Female")) + 
  theme_classic() + 
  labs(title = "There is a wide variation of expenditures by ethnicity, but not by gender",
       subtitle = "The difference by gender for White and black is deceiving\ndue to differences in age profiles and associated expenditures (see discussion)", 
       fill = "",
       x = "", 
       y = "Median expenditures", 
       tag = "",     # would likely be a figure number or something else
       caption = "\nSource: Disability.csv data set from Week 1")    # would likely be a source or reference

p
```

## Tick labels and marks

```{r}
# Change the axis ticks and labels
p <- p + 
  theme(axis.ticks.y = element_blank(), 
        axis.line.y = element_blank(),
        axis.ticks.x = element_blank()) + 
  scale_y_continuous(breaks = seq(0, 30000, 5000))   # run it from 0 --> 30000, with breaks every 5000

p
```

```{r}
# use scales to format numbers
library(scales)
p <- p + scale_y_continuous(breaks = seq(0, 30000, 5000), 
                            labels = dollar)

p
```

```{r}
# Legend location and labels
p <- p + theme(legend.position = "bottom") +
         scale_fill_manual(labels = c("Women", "Men"), values = c("#F8766D", "#00BFC4")) 

# specify the fill values, these are color codes that are the original colors to add manually
# custom labels and color for "fill = gender" was specified in aes()
# set custom axes breaks and labels for "y = med_exp" that was specified in aes()
p
```

+ Only spend the time and effort if it helps clarify your message and helps get the point across

+ Budget time in the project to work on your visuals!!

# Color

+ Color can help make something stand out or confusing your audience

```{r}
# load libraries
library(colorspace)

# Plot reset color palettes
hcl_palettes(plot = TRUE)

```

```{r}
# Basic bar graph
dis %>% 
  filter(ethnicity %in% c("Asian", "Black", "Hispanic", "White")) %>% 
  ggplot(mapping = aes(x = ethnicity)) + 
  geom_bar() + 
  coord_flip() + 
  theme_classic()
```

```{r}
# Something other than black is easier on the eyes
dis %>% 
  filter(ethnicity %in% c("Asian", "Black", "Hispanic", "White")) %>% 
  ggplot(mapping = aes(x = ethnicity)) +   # put fill = in aes() to color code the factor variables
  geom_bar(fill = "light blue") +   # color in the geom to change color of fill
  coord_flip() + 
  theme_classic()
```

```{r}
# Scatter plot
dis %>% 
  ggplot(mapping = aes(x = age, y = expenditures)) +   
  geom_point() +  
  theme_classic()
```

```{r}
# Scatter plot with color
dis %>% 
  ggplot(mapping = aes(x = age, y = expenditures)) +   
  geom_point(color = "light blue") +  
  theme_classic()
```

```{r}
# QUALITATIVE SCALE FILL
# Decorational color schemes
# Don't draw attention away from main points
dis %>% 
  filter(ethnicity %in% c("Asian", "Black", "Hispanic", "White")) %>% 
  ggplot(mapping = aes(x = ethnicity, fill = ethnicity)) + 
  geom_bar() + 
  coord_flip() + 
  theme_classic() + 
  scale_fill_discrete_qualitative(guide = "none")    # this is from colorspace()

# guide = FALSE is because there's no need for a legend since they're already labeled on the axes

```

+ Don't let color distract from main points

```{r}
# SEQUENTIAL SCALE FILL
dis %>% 
  filter(ethnicity %in% c("Asian", "Black", "Hispanic", "White")) %>% 
  ggplot(mapping = aes(x = ethnicity, fill = ethnicity)) + 
  geom_bar() + 
  coord_flip() + 
  theme_classic() + 
  scale_fill_discrete_sequential(guide = "none")
  
```

```{r}
# different geoms
dis %>% 
  ggplot(mapping = aes(x = age, y = expenditures, color = ethnicity)) + 
  geom_point() + 
  theme_classic() + 
  scale_color_discrete_sequential()

```

```{r}
# line
dis %>% 
  filter(ethnicity %in% c("Asian", "Black", "Hispanic", "White")) %>% 
  group_by(age, ethnicity) %>% 
  summarize(med_exp = median(expenditures)) %>% 
  ggplot(mapping = aes(x = age, y = med_exp, color = ethnicity)) + 
  geom_smooth(se = FALSE) + 
  theme_classic() + 
  scale_color_discrete_sequential()
```

Draw attention - make one or a set of elements stand out

```{r}
dis %>% 
  filter(ethnicity %in% c("Asian", "Black", "Hispanic", "White")) %>% 
  ggplot(mapping = aes(x = ethnicity, fill = ethnicity)) + 
  geom_bar() + 
  coord_flip() + 
  theme_classic() + 
  scale_fill_manual(values = c("gray", "gray", "gray", "red", "gray", "gray", "gray", "gray"), guide = "none")    # guide = "none" means no legend 
```

+ "fill = interaction(ethnicity, gender)" to make comparisons

    - allows using "fill" with multiple variables
    
## Qualitative

+ no individual value should stand out

```{r}
dis %>% 
  filter(ethnicity %in% c("Asian", "Black", "Hispanic", "White")) %>% 
  ggplot(mapping = aes(x = ethnicity, fill = gender)) + 
  geom_bar(position = "dodge") + 
  coord_flip() + 
  theme_classic() + 
  scale_fill_discrete_qualitative(palette = "Dark 2")    # palette came from colorspace library
```


## Sequential

+ goal is to create order

+ use "reorder()" as necessary

+ easy on the eyes and uncluttered

```{r}
dis %>% 
  filter(ethnicity %in% c("Asian", "Black", "Hispanic", "White")) %>% 
  group_by(ethnicity) %>% 
  summarise(count_n = n()) %>% 
  ggplot(mapping = aes(x = reorder(ethnicity, count_n), y = count_n, fill = ethnicity)) + 
           geom_bar(stat = "identity") +
  coord_flip() + 
  theme_classic() + 
  scale_fill_discrete_sequential(palette = "Purples 2", guide = "none")
```

## Divergence
```{r}
# Diverging color scheme with bar graph
dis %>% 
  filter(ethnicity %in% c("Asian", "Black", "Hispanic", "White")) %>% 
  ggplot(mapping = aes(x = ethnicity, fill = ethnicity)) +
  geom_bar() +
  coord_flip() +
  theme_classic() +
  scale_fill_discrete_diverging(palette = "Blue-Red 2", guide = "none")
```

```{r}
# Diverging color scheme with scatter plot
dis %>% 
  ggplot(mapping = aes(x = age, y = expenditures, color = age_group)) +
  geom_point() +
  theme_classic() +
  scale_color_discrete_diverging(palette = "Green-Orange", guide = "none")
```


# Shape, line style, and size

+ Size and shape typically used in scatter plots

    - Typically difficult to detect
    
+ Line Types typically used in line graphs

+ Size typically used on quantitative variables

+ Shape and linetype typically used on factor variables

+ Use size and shape sparingly and intentionally

## Shape

```{r}
# Basic scatter plot with base shape points

dis %>% 
  filter(ethnicity %in% c("Asian", "Black", "Hispanic", "White")) %>% 
  ggplot(mapping = aes(x = age, y = expenditures, shape = ethnicity)) + 
  geom_point() + 
  theme_classic()
```

+ Difficult to see what's going on

```{r}
# Show a list of available shapes
data.frame(shape = 0:24) %>% 
  ggplot(mapping = aes(0, 0, shape = shape)) +
  geom_point(aes(shape = shape), size = 5, fill = "red") +
  scale_shape_identity() +
  facet_wrap( ~ shape) +
  theme_void()
```


```{r}
# Basic scatter plot with different shape points and base size
dis %>% 
  filter(ethnicity %in% c("Asian", "Black", "Hispanic", "White")) %>% 
  ggplot(mapping = aes(x = age, y = expenditures, shape = ethnicity)) + 
  geom_point() + 
  theme_classic() + 
  scale_shape_manual(values = c(0, 4, 16, 3))   # set custom points
```

## Size

+ To se the base size of points, set size = #

```{r}
# Scatter plot with different shapes but larger size

dis %>% 
  filter(ethnicity %in% c("Asian", "Black", "Hispanic", "White")) %>% 
  ggplot(mapping = aes(x = age, y = expenditures, shape = ethnicity)) + 
  geom_point(size = 3) + 
  theme_classic()
```

+ To distinguish different data set size in the aes() function

```{r}
# Set size by magnitude of data
dis %>% 
  filter(ethnicity %in% c("Asian", "Black", "Hispanic", "White")) %>% 
  ggplot(mapping = aes(x = age, y = expenditures, size = expenditures)) + 
  geom_point() + 
  theme_classic()
```

+ Can't really tell the difference between the sizes

## Shape and Size

+ If you use too many different things together, you'll end up with multiple legends

```{r}
# Basic scatter plot with different size and shape points
dis %>% 
  filter(ethnicity %in% c("Asian", "Black", "Hispanic", "White")) %>% 
  ggplot(mapping = aes(x = age, y = expenditures, shape = ethnicity, size = factor(ethnicity))) +
  geom_point() +
  theme_classic() 
```

## Custom shape and size to get 1 legend

```{r}
# Basic scatter plot with different shape points
dis %>% 
  filter(ethnicity %in% c("Asian", "Black", "Hispanic", "White")) %>% 
  ggplot(mapping = aes(x = age, y = expenditures, shape = ethnicity, size = factor(ethnicity))) +
  geom_point() +
  theme_classic() +
  scale_size_manual("", values = c(4, 2, 3, 1)) +
  scale_shape_manual("", values = c(0, 4, 16, 3))
```

+ When using multiple attributes, use manual to combine multiple legends into 1

```{r}
# Basic scatter plot with different shape and color points
dis %>% 
  ggplot(mapping = aes(x = age, y = expenditures, shape = ethnicity, color = ethnicity)) +
  geom_point() +
  theme_classic()
```


```{r}
# Basic scatter plot with base shape and larger size points
dis %>%   
  ggplot(mapping = aes(x = age, y = expenditures, shape = ethnicity)) +
  geom_point(size = 2) +
  theme_classic()
```

## Linetypes

+ avoid spaghetti graphs

+ it can be difficult to decifer the base linetypes in ggplot

```{r}
# Create random dates for making line graphs
dis$date <- sample(seq(as.Date("2015/01/01"), as.Date("2018/01/01"), by = "month"), nrow(dis), replace = TRUE)
```

```{r}
dis %>% 
  filter(ethnicity %in% c("Asian", "Black", "Hispanic", "White")) %>% 
  group_by(date, ethnicity) %>% 
  summarize(med_exp = median(expenditures)) %>% 
  ggplot(mapping = aes(x = date, y = med_exp, linetype = ethnicity)) +
  geom_line() +
  theme_classic()
```


Custom Linetypes
```{r}
# Show a list of available linetypes
data.frame(linetype = factor(1:13)) %>% 
  ggplot(mapping = aes(0, 0, linetype = linetype)) +
  geom_segment(mapping = aes(x = 0, xend = 1, y = linetype, yend = linetype, linetype = linetype), size = 1.1)+
  scale_x_continuous(limits = c(0, 1), breaks = NULL, name = "") +
  scale_y_discrete(name = "") +
  guides(linetype = guide_legend(reverse = TRUE)) +
  theme_void()
```


Improving readability with linetypes

```{r}
# Custom linetypes
dis %>% 
  filter(ethnicity %in% c("Asian", "Black", "Hispanic", "White")) %>% 
  group_by(date, ethnicity) %>% 
  summarize(med_exp = median(expenditures)) %>% 
  ggplot(mapping = aes(x = date, y = med_exp, linetype = ethnicity, size = ethnicity)) +
  geom_line() +
  theme_classic() +
  scale_linetype_manual(values = c(1, 3, 6, 10)) +
  scale_size_manual(values = c(0.5, 1, 1.25, 1))
```

Use facets to show multiple line graphs

```{r}
# Using facets
dis %>% 
  filter(ethnicity %in% c("Asian", "Black", "Hispanic", "White")) %>% 
  group_by(date, ethnicity) %>% 
  summarize(med_exp = median(expenditures)) %>% 
  ggplot(mapping = aes(x = date, y = med_exp, linetype = ethnicity)) +
  geom_line() +
  theme_bw() +
  facet_wrap(~ ethnicity) +
  scale_linetype_discrete(guide = "none")
```


# Annotations

+ Trying to make a point on the visual to stand out

    - Communicate the key findings from your analysis
    
+ Easy to over kill with annotations

    - Don't obscure the point!
    
```{r}
# Base graph to start with
g_ann <- dis %>% 
  filter(ethnicity %in% c("Asian", "Black", "Hispanic", "White")) %>% 
  group_by(ethnicity, gender) %>% 
  summarize(med_exp = median(expenditures)) %>% 
  ggplot(mapping = aes(x = reorder(ethnicity, med_exp), y = med_exp, fill = gender)) + 
  geom_bar(stat = "identity", position = "dodge") + 
  coord_flip() + 
  scale_fill_discrete(limits = c("Male", "Female")) + 
  theme_classic()

g_ann
```

## Annotation as context

```{r}
# Annotation as context
# Showing how much median expenditures of each element is vs. the median

g_ann + 
  geom_hline(yintercept = median(dis$expenditures)) +
  annotate(geom = "text", x = "Hispanic", 
           y = 200 + median(dis$expenditures),
           label = "Median expenditures", hjust = 0)

```


```{r}
# Shading to highlight a finding
g_ann + 
  annotate(geom = "rect", xmin = 3.5, xmax = 4.5,
           ymin = -200, ymax = 200 + median(dis$expenditures[dis$ethnicity == "White" & dis$gender == "Male"]),
           fill = "lightblue", alpha = 0.5) +
  annotate(geom = "rect", xmin = 1.5, xmax = 2.5,
           ymin = -200, ymax = 200 + median(dis$expenditures[dis$ethnicity == "Black" & dis$gender == "Male"]),
           fill = "lightblue", alpha = 0.5) 

```

```{r}
# Create random dates for making line graphs
set.seed(123)
dis$date <- sample(seq(as.Date("2015/01/01"), as.Date("2018/01/01"), by = "month"), nrow(dis), replace = TRUE)
```


```{r}
# Shading of time series
# Look at median expenditures over time 
dis %>% 
  group_by(date) %>% 
  summarize(med_exp = median(expenditures)) %>% 
  ggplot(mapping = aes(x = date, y = med_exp)) + 
  geom_line() +
  annotate(geom = "rect", xmin = as.Date("2015-05-15"), 
           xmax = as.Date("2015-08-15"), 
           ymin = 3000, ymax = 33000, 
           fill = "lightblue", alpha = 0.5) +
  theme_classic()
```

+ Use annotations to set the context to highlight the finding within the bigger picture

