---
title: "week_04_notes"
author: "Andrew Nalundasan"
date: "7/9/2021"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggplot2)
library(vtable)
library(grid)
library(gridExtra)
library(lattice)
library(janitor)
library(GGally)
```

# Overview Video

+ Focus on getting good at ggplot2

+ Data Vizualization checklist

    - URL
    
+ pie charts

    - limited value
    - difficult to tell relative sizes of different categories
    - works fine with only 2 categories
    - 3 categories is still difficult to tell a good story
    
+ tree map is a good alternative to pie chart

    - instead of a pie, it's squares and blocks
    - library(treemapify)
    - geom_treemap()
   
+ scales functions

    - set limits of the graph
    - bar graph: 
        - limits = c("C", "A", "B")
        - this makes the bars in C A B order

+ labels

    - can use unnamed functions
    - function(x) or \(x)
    
+ theme

    - strip 
        - useful when using facets to label the different facets
    - title
    - element
        - element text, element rectangles, element blank, etc.
    - geometry
        - if aesthetic is INSIDE the aesthetic function, it applies to just the one element
        - if aesthetic is OUTSIDE the aesthetic function, it applies to the whole viz
    - font size is determined differently between theme vs geometry
    - theme is pretty standard how we know it
    - geometry you must divide by .pt
        - size = 12/.pt
        - this gives standard 12 point font
        - conversion is also necessary in annotate() functions

+ guides
    
    - color = 'none' to remove legends
    
```{r}
# knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

# this will remove messages and warnings from R outputs. 
```


+ annotation

    - prefer annotation over legends (where possible)
        - move the annotation closer to the data
    - theme(legend.position) can move the legend closer to the data
    
    
# Best Practices



# Clutter is your enemy

# Gestalt principles

# Visual order, part 1

# Visual order, part 2

# Focusing audience attention

# Titles, labels, and legends

# Color

# Shape, line style, and size

# Annotations
